Lexical frequency mediates the facilitation effect of cognateness in bilingual lexical acquisition

### 2023-03-29 manuscript/appendix/_model-details.qmd
New version of the manuscript, with revised Discussion

To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($Age$), word-form length ($Length$), word exposure index ($Exposure$), and of phonological similarity ($Levenshtein$), the two-way interactions $Age \times Exposure$, $Age \times Levenshtein$, and $Exposure \times Leveshtein$, and the three-way interaction $Age \times Exposure \times Leveshtein$. We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression]. For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].

$$
\begin{aligned}
\textbf{Linear model:} \\
logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Length}_{ij} + & \\
& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
& (\beta_{4} + u_{4_{i}}) · \text{Levenshtein}_{ij} + & \\
& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Levenshtein}_{ij}) + & \\
& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Levenshtein}_{ij}) & \\
\end{aligned}
$$ {#eq-linear}

@@

We used the Bayesian framework to estimate the parameters in our model. This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical]. This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations. We used a weakly informative prior for our parameters, with the exception of the main effect of $Age$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].

>>>>

To test our hypotheses, we included several predictors in the regression model as fixed effects: the main effects of participant age ($\text{Age}$), word-form length ($\text{Length}$), word exposure index ($\text{Exposure}$), and of phonological similarity ($\text{Cognateness}$), the two-way interactions $\text{Age} \times \text{Exposure}$, $\text{Age} \times \text{Cognateness}$, and $\text{Exposure} \times text{Leveshtein}$, and the three-way interaction $\text{Age} \times \text{Exposure} \times \text{Leveshtein}$. We also included crossed random effects for participants and translation equivalents to account for the repeated measures in our dataset---each participant provided responses to multiple translation equivalents, and each translation equivalent was responded to by multiple participants [@gelman2020regression]. For both grouping variables, we included random intercepts, random slopes, and correlation parameters for all predictors were repeated measures were observed in our dataset [@barr2013random, see @eq-linear].

$$
\begin{aligned}
\textbf{Linear model:} \\
logit(p_{k}) = \text{ln} \frac{p_{k}}{1-p_{k}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + \\
& (\beta_{1} + u_{1_{i}} + w_{1_{j}}) · \text{Age}_{i} + & \\
& (\beta_{2} + u_{2_{i}} + w_{2_{j}}) · \text{Length}_{ij} + & \\
& (\beta_{3} + u_{3_{i}} + w_{3_{j}}) · \text{Exposure}_{ij} + & \\
& (\beta_{4} + u_{4_{i}}) · \text{Cognateness}_{ij} + & \\
& (\beta_{5} + u_{5_{i}} + w_{3_{j}}) · (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
& (\beta_{6} + u_{6_{i}}) · (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
& (\beta_{7} + u_{7_{i}}) · (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
& (\beta_{8} + u_{8_{i}}) · (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
\end{aligned}
$$ {#eq-linear}

@@

We used the Bayesian framework to estimate the parameters in our model. This involves using the Bayes theorem to compute a distribution (*posterior distribution*) that describes what values of each parameter in the model are more likely given the data (*likelihood*, see @eq-likelihood), and previous knowledge about such distribution (*prior distribution*, see @eq-prior) [@mcelreath2020statistical]. This posterior distribution not only informs about the most likely values of our regression coefficients of interest, but also about the uncertainty around such estimations. We used a weakly informative prior for our parameters, with the exception of the main effect of $\text{Age}$, for which we specified a strongly informative prior based on previous literature about how age affects the acquisition of words [see @eq-prior].

### 2023-03-29 manuscript/appendix/_model-diagnostics.qmd
New version of the manuscript, with revised Discussion

We used Stan [@carpenter2017stan] as the probabilistic language behind the estimation of our Bayesian models in this study, with `brms` as its R interface [@burkner2017brms]. This language implements the Markov Chain Monte-Carlo (MCMC) algorithm to explore the posterior distribution of the model. Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined. We run `r dim(model_fit_4$fit)[2]` MCMC chains, each `r dim(model_fit_4$fit)[1]` iterations long each. The correct performance of this algorithm is critical to the quality of the statistical evidence to which the outcomes of the model lead.

@@

```{r}
#| label: fig-model-rhats-neffs
#| fig-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6
#| fig-width: 9
map_df(model_rhats, bind_rows, .id = "model") |> 
    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
    pivot_longer(-model, 
                 names_to = ".variable",
                 values_to = ".rhat") |>
    filter(str_detect(.variable, "b_|sd_")) |> 
    drop_na() |> 
    ggplot(aes(.rhat, colour = model, fill = model)) +
    facet_wrap(~model, nrow = 1) +
    geom_vline(xintercept = 1.01, colour = "grey", linewidth = 0.75) +
    geom_histogram(bins = 12, colour = "white") +
    geom_rug(alpha = 0.5) +
    labs(x = "Gelman Rubin diagnostic (R-hat)",
         y = "MCMC samples",
         title = "Gelman-Rubin (R-hat) diagnostic") +

@@

## Posterior draws and bi-variate scatterplots

```{r}
#| label: fig-model-pairs
#| fig-cap: "Marginal distribution and bi-variate scatterplot of posterior samples for the fixed regression coefficients in Model 4."
#| fig-height: 12
#| fig-width: 14
#| echo: false
#| warning: false
#| message: false
titles <- c("Comprehension",
            "Production",
            "Age",
            "Phonemes",
            "Exposure",
            "Levenshtein",
            "Age × Exposure",
            "Age × Levenshtein",
            "Exposure × Levenshtein",
            "Age × Exposure × Levenshtein")

posterior <- as.array(model_fit_4)
np <- nuts_params(model_fit_4) 
mcmc_pairs(posterior,
           np = np, 
           pars = c("b_Intercept[1]",
                    "b_Intercept[2]",
                    "b_age_std",
                    "b_n_phon_std",
                    "b_exposure_std",
                    "b_lv_std",
                    "b_age_std:exposure_std",
                    "b_age_std:lv_std",
                    "b_exposure_std:lv_std",
                    "b_age_std:exposure_std:lv_std"),
           grid_args = list(subtitles = titles),
           off_diag_args = list(size = 0.5,
                                alpha = 1/10))
```

## Posterior-predictive checks

```{r}
#| label: fig-model-ppc
#| fig-cap: "Model posterior predictive checks (PPC). Bars indicate the observed proportion of responses to each category (*No*, *Understands*, and *Understands and Says*). Blue dots and error bars represent the mean proportion of responses simulated from the posterior for each category, and its 95% interval."
#| eval: true
#| echo: false
#| warning: false
#| message: false
y_int <- as.integer(as.integer(responses$response))
color_scheme_set(c(clrs[c(4, 4, 3, 2, 1)], "white"))

>>>>

We used Stan [@carpenter2017stan] as the probabilistic language behind the estimation of our Bayesian models in this study, with `brms` as its R interface [@burkner2017brms]. This language implements the Markov Chain Monte-Carlo (MCMC) algorithm to explore the posterior distribution of the model. Broadly, this algorithm is used to iteratively sample the joint sampling space of the parameters to be estimated in the model, and compute, for each value sampled, its likelihood under some probability distribution previously defined. We run `r dim(model_fit_3$fit)[2]` MCMC chains, each `r dim(model_fit_3$fit)[1]` iterations long each. The correct performance of this algorithm is critical to the quality of the statistical evidence to which the outcomes of the model lead.

@@

```{r}
#| label: fig-model-rhats-neffs
#| fig-cap: "Gelman-Rubin (R-hat) convergence diagnostic of the MCMC chains in the Bayesian models."
#| echo: false
#| message: false
#| warning: false
#| fig-height: 4
#| fig-width: 9
map_df(model_rhats, bind_rows, .id = "model") |> 
    mutate(model = paste0("Model ", str_extract(model, "[0-9]"))) |> 
    pivot_longer(-model, 
                 names_to = ".variable",
                 values_to = ".rhat") |>
    filter(str_detect(.variable, "b_|sd_")) |> 
    drop_na() |> 
    ggplot(aes(.rhat, colour = model, fill = model)) +
    facet_wrap(~model, nrow = 1) +
    geom_vline(xintercept = 1.01, colour = "grey", linewidth = 0.75) +
    geom_histogram(bins = 12, colour = "white") +
    geom_rug(alpha = 0.5) +
    labs(x = "Gelman Rubin diagnostic (R-hat)",
         y = "MCMC samples",
         title = "Gelman-Rubin (R-hat) diagnostic") +

@@

## Posterior draws and bi-variate scatterplots {.unnumbered}

```{r}
#| label: fig-model-pairs
#| fig-cap: "Marginal distribution and bi-variate scatterplot of posterior samples for the fixed regression coefficients in Model 4."
#| fig-height: 12
#| fig-width: 14
#| echo: false
#| warning: false
#| message: false
clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

color_scheme_set(rep(clrs[4], 6))

titles <- c("Comprehension",
            "Production",
            "Age",
            "Length",
            "Exposure",
            "Cognateness",
            "Age × Exposure",
            "Age × Cognateness",
            "Exposure × Cognateness")

posterior <- as.array(model_fit_3)
np <- nuts_params(model_fit_3) 
mcmc_pairs(posterior,
           np = np, 
           pars = c("b_Intercept[1]",
                    "b_Intercept[2]",
                    "b_age_std",
                    "b_n_phon_std",
                    "b_exposure_std",
                    "b_lv_std",
                    "b_age_std:exposure_std",
                    "b_exposure_std:lv_std"),
           grid_args = list(subtitles = titles),
           diag_args = list(fill = clrs[1]),
           off_diag_args = list(size = 0.5,
                                alpha = 1/10))
```

## Posterior-predictive checks {.unnumbered}

```{r}
#| label: fig-model-ppc
#| fig-cap: "Model posterior predictive checks (PPC). Bars indicate the observed proportion of responses to each category (*No*, *Understands*, and *Understands and Says*). Blue dots and error bars represent the mean proportion of responses simulated from the posterior for each category, and its 95% interval."
#| eval: true
#| echo: false
#| warning: false
#| message: false
y_int <- as.integer(as.integer(responses$response))
color_scheme_set(c(clrs[c(4, 4, 2, 2, 2)], "white"))
